Plant species traits and growth rates: meta-analysis shows correlations change with plant size as predicted

### 2015-08-10 MS.tex
Add 14 quality criteria for meta-analyses - add table - appendix D - ref

\newpage

@@

\textit{Type of article} Letters

\textit{Number of words in the abstract} 150

\textit{Number of words in the main text} 4583

\textit{Number of words in each text box} 770

\textit{Number of general references} 50

\textit{Number of data source references} 112

\textit{Number of figures} 2

\textit{Number of tables} 1

\textit{Number of text boxes} 1

@@

Meanwhile the possibility has been raised that trait-growth correlations may change systematically with size \citep{Enquist:2007ek,Falster:2011ii, Ruger:2012jv, Iida:2014ep, Iida:2014hq}. Recent theory suggests that the absence of a correlation between SLA and growth at larger sizes may be expected, when potential influences of traits on plant function are considered \citep{Falster:2011ii, falster:2013}. At larger sizes the low-cost deployment benefits of high SLA may be offset or even outweighed by the high costs of rapid leaf turnover (see Text Box 1\ref{sec:growth}). Today, we have no

@@

\textbf{Effect on absolute (AbsGR) and relative growth rate (RGR}
Tree growth is the increase in dimensions (i.e. diameter, height or mass) of an individual tree through time. While plant functional traits measurements are standardized across plant ontogeny, plant growth is commonly expressed differently according to the plant size/stage considered. Thus, mass and height are both used for smaller plants, whereas height and diamater are more convenient measures for larger tree. In addition, the growth increment (i.e. rate of change of a dimension per unit time) can be expressed in either absolute or relative terms. Relative growth rate (RGR) is commonly used for smaller plants, but is less useful for larger-plants because there is a very rapid decline in RGR with increasing tree size due to accumulation of support tissues (bark, branches etc). According to the growth model used here, the correlation between trait and growth is predicted to revert according to the growth measurement used for both SLA and seed mass (see Text Box 1\ref{sec:growth}. The predicted relationship between growth and either SLA or seed mass in seedlings differs if absolute growth rate in mass is used instead of $RGR_{mass}$ (i.e. becomes negative for SLA, positive for SM).

@@

We acknowledge that it would be better to address plant size as a continous variable, but the limited number of correlations available in the literature along plant size constrains us to approximate plant size by plant stage. The authors of the 112 articles reported in our meta-analyses, categorized plants into four differents stages (i.e. seedling, juvenile, sapling, adult). They also reported plant size informations in diameter, height or age; generally the authors give the size of the smallest and the biggest plants species used in their analyses. While the stage categories have been established according to the expertise and the appreciation of the authors, by compiling data across multiple studies and across different growth form (tree, woody, etc), there is a risk of assigning plants of similar size to different categories. Thus, we tested how consistent were the range values in size/age attributed to each stage across the 112 studies, and across the different growth form (tree, woody and across growth form, Fig. A1). Suprisingly, the threshold in size/age used by the authors to defined a plant stage are quite steady across the 112 studies (Fig. A1). In particular, there was a clear break between seedlings and saplings in both age (at 1 yr) and height (at 0.5 m); there was also a clear break between saplings and adults in diameter (at 10 cm). The distinctions between sapling-adult in age (around 5 yrs) and height (around 2 m), and between seedling-sapling in diameter (around 1cm) were less pronounced. We used simple allometric relationships from the Biomass And Allometry Database (BAAD) \citep{Falster:2015} to ensure that these size/stage cutoffs match across the different size measurements (i.e. diameter, height and age, Fig. A2). For instance, we showed according to BAAD that a sapling with a diameter < 10 cm would have a height < 1 m. Thus, according to the size/stage threshold established across 112 studies, the plant will be assigned to the sapling stage whatever we focused on its diameter or height (see Fig. A2 c).

There were an ambiguity on the plant stage attribution for less than 5\% of the studies (i.e. 5 studies on 112). We re-red carefully each of these studies, looking at any particular reason addressed by authors that could explained these discrepancies. Interestingly, these discrepancies were not due to a differences between growth form; all the ambiguous cases corresponds to studies focusing exclusively on tree species (Fig. A1). In fact, we identified two main explanations. First, the term "juvenile" was vague enough to cover either seedling or sapling stages according to the authors. Second, the reproductive state of a plant is not strictly correlated to plant size; some plants with a large stature are not assessed as adult because the authors selected them to be non reproductive (See ID 50, 52 and 105 in Fig. A1). Thus, we attributed a study to a new stage category when 95\% of the recorded range of sizes for that study lay within a new category. And, we attributed a study to the adult stage when the authors identified the plants as reproductive regardless the plant size or age. The size/stage threshold defined across the 112 articles, allowed us to attribute straightforwardly a stage category to studies for which no stage have been distinctly mentionned by the authors (5\% of the data, noted as 'mix' in Fig. A1). Finally, we run the analyses with both the old and the standardized categories to ensure that the different stage attribution will not disturb our results.

Results using the categories established by the authors are similar than the one using the standardized categories and are available in Fig. A8 model.1. In the main text, we report the results using the standardized categories. In the rest of the mansucript, we preferred the term "juvenile" to "seedling", because for a portion of the scientific community "seedling" also refers to a particular experimental condition (a greenhouse experiment).

@@

We tested stage effect on traits for which > 15 correlation coefficients across at least two stages were recorded. We used a mixed-effect model of meta-analysis, in which it is assumed that studies within a stage share a common mean effect but that there is also random variation among studies in a stage, in addition to sampling variation. This random effect allows us to account for the heterogeneity among studies (e.g.~experimental type, vegetation type, error, etc.). For each trait, we established if stage could explain variance in the coefficient correlation by comparing models with the stage variable to the null model using log-likelihood ratios \citep{Zuur:2009cfa}. If the log-likelihood test returned a significant p-value, the stage model fitted the data better than the null model. Estimates of the effect size for each stage were considered significant if the 95\% confidence intervals did not overlap with zero. In the case of multiple comparisons within a study, we used the method proposed by \citet{Borenstein:2009um}: we calculated the mean correlation coefficient by trait and by study, and re-calculated the effect size.

Following recent recommendations for the adoption of a checklist of quality criteria for meta-analysis in ecology (\citealt{Koricheva:2014ku}), we conducted several analyses to assess the reliability of our meta-analysis. Table A2 in Appendix 1 reported how and where (i.e in the text, appendix and/or figures) we addressed the 14 methodological quality criteria listed by \citet{Koricheva:2014ku}.

@@

For both SLA and SM, the correlation between growth rate and the trait changed significantly with plant size in the manner expected. Model fits improved when stage was incorporated (see log likelihood ratio tests, Fig. \ref{Fig2} a and d; average values for the coefficients r ($r_{\textrm{mean}}$) in Fig. A5). The correlation between growth and SLA shifted from strongly positive at juvenile stage $r_{\textrm{mean}}=0.6 \pm 0.08$, Mean +SD) to not significantly different from zero at adult stage ($r_{\textrm{mean}}= -0.01 \pm 0.07$, CI 95\% overlapping zero in Fig. \ref{Fig2} a). The correlation between growth and SM shifted from negative at juvenile stage ($r_{\textrm{mean}}= -0.59 \pm 0.07$) to not significantly different from zero at adult stage ($r_{\textrm{mean}}= -0.09 \pm 0.02$, , CI 95\% overlapping zero in Fig. \ref{Fig2} d). For both traits,  correlation coefficients recorded at sapling stage were intermediate between juvenile and adult stages.

For WD, Hmax and Aarea, correlations with growth were also reasonably consistent with expectations, though with limitations arising from very unequal numbers of observations between growth stages. For Hmax, the expectation was for a correlation with growth to be absent for small plants but present at intermediate and larger sizes, the mechanism being increased relative allocation to reproduction as Hmax is approached. Indeed Hmax was positively correlated with growth at adult stage ($r_{\textrm{mean}}= 0.38 \pm 0.05$, n=18). At sapling and juvenile stages, the correlation between Hmax and growth was not significantly different from that observed at adult stage (model fit not improved when stage was incorporated, see log likelihood ratio tests in Fig. \ref{Fig2}c), but also was consistent with no correlation (Fig \ref{Fig2}c, confidence intervals overlapping zero), due to much smaller numbers of replicate studies. Similarly for WD and for Aarea, results were consistent with negative (WD) and positive (Aarea) correlations with growth across all plant sizes (Fig 2 b and e, model fit not improved when stage was incorporated). However, at the same time confidence intervals of r often spanned zero especially where there were small numbers of studies for particular growth stages.

@@

Given our basic conclusion that reported correlations responded to plant size in a manner consistent with expectations from theory, we assess now the coverage of the available data, to determine if biased sampling could have erroneously generated the above results. The results reported were confined to correlations across species, and under non-stressed conditions (n=176). A "complete" dataset (n=216) also included correlations measured under stressed conditions, and spanning different plant stages between the growth and trait measurements. This "complete" dataset generated similar effect sizes to those reported above from the "ideal" dataset, and led to similar conclusions (Fig. \ref{Fig2} grey points). Differences concerned often the confidence intervals around particular mean effect sizes in Fig. \ref{Fig2}. These differences in CI were not easily attributed to any single cause. For instance, the correlation between growth and WD at juvenile stage shifted from negative to non significant (i.e CI95\% included zero) due to 2 positive correlation coefficients from \citet{Augspurger:1984ct};  these correlations unlike others used wood density values from adult plants together with growth from juveniles. For Aarea, the comparison between complete and ideal data showed no signal of an effect of stress (here mainly low light) on the correlation with growth at seedling stage; the correlation was still positive between Aarea and growth even when 7 correlations measured under low light conditions were added. Finally, our results were unlikely to be biased by dominance of one or a few studies that measured a large number of responses, since when multiple responses were condensed into a single average within a trait and within a study, outcomes were very similar (Fig. A7).

We were pleasantly surprised to find that correlations had been reported for at least two different stages across 18 traits out of the 20 targeted (Fig. \ref{Fig1}). We had thought it likely that for some traits correlations would be reported only for one size class, or only mixed across different sizes. These limitations applied only for LAR and NARmass, where studies appeared confined to juvenile stages. However, among the 20 traits studied we had enough data to test the effect of stage on the trauts value only for the five for which we had predictions.

We found 112 studies published between between 1983 and 2014 that met our requirements. Climate zones spanned from temperate regions to  tropics (map in Fig. A3). Different experiment types (laboratory, field, plantation or greenhouse), growth measurements and growth form (tree, woody or across growth form) were included. Given this variety, it was not surprising to find wide ranges of r reported for all traits. For instance, r varied from -0.98 to 0.99 for SLA, and from -0.74 to 0.74 for WD (Fig. A4). The growth form, the experiment types and the growth measurement were not markedly best predictors of the data (results shown in the Fig. A8). Number of species used for each correlation ranged from 2 to 300 species with average 31.

Finally, we met all the quality criteria required for meta-analysis in ecology. We found no evidence of publication bias (). The extend of the heterogenity was high in our meta-analysis, and substantially explained by the plant stage for SLA and Seedmass as expected. Finally, we showed no temporal change in effect size for all the traits.

@@

For seedlings our meta-analyses showed as expected a negative relationship between seed mass and relative growth rate (in mass or height or volume) that disappeared across plant stage. This negative relationship arises because seedlings from larger seeds are larger to begin with, as extensively discussed in the literature \citep[reviewed by][]{Turnbull:2012ew}. It applies only to RGR not to absolute growth rate, but for seedlings growth is most often reported as RGR (Appendix FigA10 d).

@@

\newpage
\section*{Figures}

\begin{figure}[htbp]
\centering
\includegraphics{output/Fig1.pdf}
\caption{Number of correlations between trait and growth for each plant stage recorded in our dataset. Shown here are 'raw dataset' (recorded both across and within species, under unstressed and stressed conditions). Bar colors: red for adult, orange for sapling, grey for juvenile. Aarea (or Amass) = rate of $CO_{2}$ assimilation per unit leaf area (or leaf mass) = maximum photosynthetic rate, Hmax= asymptotic maximum height,  Ks =stem hydraulic conductance, LA = leaf area, LMR = leaf mass ratio, NARarea (or NARmass) = net mass assimilation rate per unit of leaf area (or leaf mass), Nmass (or Narea) = nitrogen content per unit of leaf mass (or leaf area), Pmass = phosphorus content per unit of leaf mass, SA/LA = sapwood area per leaf area, SLA = specific leaf area, thickness = leaf thickness, Vessel size = stem conduit size, vessel density = stem conduit density, WD = wood density}
\label{Fig1}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{output/Fig2.pdf}
\caption{The relationship between growth rates and SLA or seed mass changes with plant stage. Plots show the effect size (+ 95\%CI) of the correlation across species between growth rate and five functional traits for plant stage comparisons. \textbf{a)} Specific leaf area (SLA), \textbf{b)} Wood density (WD), \textbf{c)} Asymptotic height (Hmax) \textbf{d)} Seed mass and \textbf{e)} Assimilation rate of CO2 (Aarea). Effect size is a standardized measure of the deviation of correlation coefficient from zero (Fisher's z-transformed). Effects are significant if confidence intervals (95\%CI) for a given stage do not overlap with zero. Shown here are 'across species dataset' (recorded only across species, and under both unstressed and stressed conditions). Black point: unstressed conditions, grey points: all the data. 'n' indicates number of correlations contributing to each effect. Likelihood Ratio Tests (LRT, for unstressed conditions only) compared maximum-likelihood fits between a pooled model and one separated by stage. A significant p-value means the stage model fitted the data better than the null model. 
}
\label{Fig2}
\end{figure}

\newpage
\section*{Table} \label{sec:table}

\begin{table}[ht]
\centering
\caption{\textbf{ Hypothesised effects of traits on key elements of plant function determining growth rate.} Arrows indicate the effect an increase in trait value would have on the main elements of growth eqs. \ref{eq:eq0}-\ref{eq:eq4} in Text box 1:  Leaf deployment per mass ($\frac{\textrm{d}a_\textrm{l}}{\textrm{d}m_\textrm{t}}$), Allocation to vegetative growth ($\frac{\textrm{d}m_\textrm{t}}{\textrm{d}b}$), Net live biomass production ($\frac{\textrm{d}b}{\textrm{d}t}$), Sapwood turnover ($k_s a_s$). SLA: Specific leaf area, WD: Wood density, Hmax: Asymptotic maximum height, SM: seed mass, Aarea: Maximum photosynthetic rate. For further details, see Text Box 1.}
\vspace{1cm}
  \begin{tabular}{l p{3cm} p{2cm}p{4cm} p{2cm}}
  & Leaf deployment \newline per mass
  & Allocation to vegetative growth
  & Net live biomass \newline production
  & Sapwood turnover\\
  & & & &\\\hline
  & & & & \\
Increase in : & & & & \\ 
  SLA &$\nearrow$ &$\_$ & $\searrow$ \newline (via $\nearrow$ of tissue turnover) &$\_$ \\
  WD & $\searrow$ &$\_$  &$\_$ &$\_$ \\
  Hmax &$\_$ &$\nearrow$ &$\_$ &$\_$ \\
  SM &$\_$&$\_$&$\_$& $\_$\\ 
  Aarea &$\_$ &$\_$ & $\nearrow$ &$\_$ \\

\hline
  \end{tabular}
\label{tab:trade-offs}
\end{table}

\newpage
\section*{Text Box 1} \label{sec:growth}

@@

\newpage
\begin{appendices}\label{sec:appendices}
\renewcommand{\thefigure}{S\arabic{figure}}
\renewcommand{\thetable}{S\arabic{table}}

@@

\newpage
\section{Supporting tables}\label{app:supp_info_tables}

>>>>

\clearpage

@@

@@

Meanwhile the possibility has been raised that trait-growth correlations may change systematically with size \citep{Enquist:2007ek,Falster:2011ii, Ruger:2012jv, Iida:2014ep, Iida:2014hq}. Recent theory suggests that the absence of a correlation between SLA and growth at larger sizes may be expected, when potential influences of traits on plant function are considered \citep{Falster:2011ii, falster:2013}. At larger sizes the low-cost deployment benefits of high SLA may be offset or even outweighed by the high costs of rapid leaf turnover (see Text Box 1\ref{sec:growth}).

@@

\textbf{Effect on absolute (AbsGR) and relative growth rate (RGR)}

The correlation between trait and growth is predicted to revert according to the growth measurement used for both SLA and seed mass (see Text Box 1\ref{sec:growth}). Tree growth in diameter, height or mass can be expressed in either absolute (AbsGR) or relative terms (RGR). Here, the predicted relationship between growth and either SLA or seed mass in seedlings differs if absolute growth rate in mass is used instead of $RGR_{mass}$ (i.e. becomes negative for SLA, positive for SM).

@@

We acknowledge that it would be better to address plant size as a continous variable, but the limited number of correlations available in the literature along plant size constrains us to approximate plant size by plant stage. The authors of the 112 articles reported in our meta-analyses, categorized plants into four differents stages (i.e. seedling, juvenile, sapling, adult). They also reported plant size informations in diameter, height or age (generally the authors give the size of the smallest and the biggest plants species used in their analyses).

While the stage categories have been established according to the expertise and the appreciation of the authors, by compiling data across multiple studies and across different growth form (tree, woody, etc), there is a risk of assigning plants of similar size to different categories. Thus, we tested how consistent were the range values in size/age attributed to each stage across the 112 studies, and across the different growth form (tree, woody and across growth form, (Fig. \ref{Fig2}). Suprisingly, the threshold in size/age used by the authors to defined a plant stage are quite steady across the 112 studies (Fig. \ref{Fig2}). In particular, there was a clear break between seedlings and saplings in both age (at 1 yr) and height (at 0.5 m); there was also a clear break between saplings and adults in diameter (at 10 cm). The distinctions between sapling-adult in age (around 5 yrs) and height (around 2 m), and between seedling-sapling in diameter (around 1cm) were less pronounced. We used simple allometric relationships from the Biomass And Allometry Database (BAAD) \citep{Falster:2015} to ensure that these size/stage cutoffs match across the different size measurements (i.e. diameter, height and age, Fig. \ref{FigA1}). For instance, we showed according to BAAD that a sapling with a diameter < 10 cm would have a height < 1 m. Thus, according to the size/stage threshold established across 112 studies, the plant will be assigned to the sapling stage whatever we focused on its diameter or height (see Fig. \ref{FigA1} c).

There were an ambiguity on the plant stage attribution for less than 5\% of the studies (i.e. 5 studies on 112). We re-red carefully each of these studies, looking at any particular reason addressed by authors that could explained these discrepancies. Interestingly, these discrepancies were not due to a differences between growth form; all the ambiguous cases corresponds to studies focusing exclusively on tree species (Fig. \ref{FigA1}). In fact, we identified two main explanations. First, the term "juvenile" was vague enough to cover either seedling or sapling stages according to the authors. Second, the reproductive state of a plant is not strictly correlated to plant size; some plants with a large stature are not assessed as adult because the authors selected them to be non reproductive (See ID 50, 52 and 105 in Fig. \ref{FigA1}). Thus, we attributed a study to a new stage category when 95\% of the recorded range of sizes for that study lay within a new category. And, we attributed a study to the adult stage when the authors identified the plants as reproductive regardless the plant size or age. The size/stage threshold defined across the 112 articles, allowed us to attribute straightforwardly a stage category to studies for which no stage have been distinctly mentionned by the authors (5\% of the data, noted as 'mix' in Fig. \ref{Fig2}). Finally, we run the analyses with both the old and the standardized categories to ensure that the different stage attribution will not disturb our results.

Results using the categories established by the authors are similar than the one using the standardized categories and are available in Fig. \ref{FigA5} model.1. In the main text, we report the results using the standardized categories. In the rest of the mansucript, we preferred the term "juvenile" to "seedling", because for a portion of the scientific community "seedling" also refers to a particular experimental condition (a greenhouse experiment).

@@

We tested stage effect on traits for which > 15 correlation coefficients across at least two stages were recorded. We used a mixed-effect model of meta-analysis, in which it is assumed that studies within a stage share a common mean effect but that there is also random variation among studies in a stage, in addition to sampling variation. This random effect allows us to account for the heterogeneity among studies (e.g.~experimental type, vegetation type, error, etc.). For each trait, we established if stage could explain variance in the coefficient correlation by comparing models with the stage variable to the null model using log-likelihood ratios \citep{Zuur:2009cfa}. If the log-likelihood test returned a significant p-value, the stage model fitted the data better than the null model. Estimates of the effect size for each stage were considered significant if the 95\% confidence intervals did not overlap with zero. In the case of multiple comparisons within a study, we used the method proposed by \citet{Borenstein:2009um}: we calculated the mean correlation coefficient by trait and by study, and re-calculated the effect size. Following recent recommendations for the adoption of a checklist of quality criteria for meta-analysis in ecology (\citealt{Koricheva:2014ku}), we conducted several analyses to assess the reliability of our meta-analysis. Table A2 in Appendix 1 reported how and where (i.e in the text, appendix and/or figures) we addressed the 14 methodological quality criteria listed by \citet{Koricheva:2014ku}.

@@

For both SLA and SM, the correlation between growth rate and the trait changed significantly with plant size in the manner expected. Model fits improved when stage was incorporated (see log likelihood ratio tests, Fig. \ref{Fig3} a and d; average values for the coefficients r ($r_{\textrm{mean}}$) in Fig. \ref{FigA4}). The correlation between growth and SLA shifted from strongly positive at juvenile stage $r_{\textrm{mean}}=0.6 \pm 0.08$, Mean +SD) to not significantly different from zero at adult stage ($r_{\textrm{mean}}= -0.01 \pm 0.07$, CI 95\% overlapping zero in Fig. \ref{Fig3} a). The correlation between growth and SM shifted from negative at juvenile stage ($r_{\textrm{mean}}= -0.59 \pm 0.07$) to not significantly different from zero at adult stage ($r_{\textrm{mean}}= -0.09 \pm 0.02$, , CI 95\% overlapping zero in Fig. \ref{Fig3} d). For both traits, correlation coefficients recorded at sapling stage were intermediate between juvenile and adult stages.

For WD, Hmax and Aarea, correlations with growth were also reasonably consistent with expectations, though with limitations arising from very unequal numbers of observations between growth stages. For Hmax, the expectation was for a correlation with growth to be absent for small plants but present at intermediate and larger sizes, the mechanism being increased relative allocation to reproduction as Hmax is approached. Indeed Hmax was positively correlated with growth at adult stage ($r_{\textrm{mean}}= 0.38 \pm 0.05$, n=18). At sapling and juvenile stages, the correlation between Hmax and growth was not significantly different from that observed at adult stage (model fit not improved when stage was incorporated, see log likelihood ratio tests in Fig. \ref{Fig3} c), but also was consistent with no correlation (Fig \ref{Fig3} c, confidence intervals overlapping zero), due to much smaller numbers of replicate studies. Similarly for WD and for Aarea, results were consistent with negative (WD) and positive (Aarea) correlations with growth across all plant sizes (Fig. \ref{Fig3} b and e, model fit not improved when stage was incorporated). However, at the same time confidence intervals of r often spanned zero especially where there were small numbers of studies for particular growth stages.

@@

Given our basic conclusion that reported correlations responded to plant size in a manner consistent with expectations from theory, we assess now the coverage of the available data, to determine if biased sampling could have erroneously generated the above results. The results reported were confined to correlations across species, and under non-stressed conditions (n=176). A "complete" dataset (n=216) also included correlations measured under stressed conditions, and spanning different plant stages between the growth and trait measurements. This "complete" dataset generated similar effect sizes to those reported above from the "ideal" dataset, and led to similar conclusions (Fig. \ref{Fig3} grey points). Differences concerned often the confidence intervals around particular mean effect sizes in Fig. \ref{Fig3}. These differences in CI were not easily attributed to any single cause. For instance, the correlation between growth and WD at juvenile stage shifted from negative to non significant (i.e CI95\% included zero) due to 2 positive correlation coefficients from \citet{Augspurger:1984ct};  these correlations unlike others used wood density values from adult plants together with growth from juveniles. For Aarea, the comparison between complete and ideal data showed no signal of an effect of stress (here mainly low light) on the correlation with growth at seedling stage; the correlation was still positive between Aarea and growth even when 7 correlations measured under low light conditions were added. Finally, our results were unlikely to be biased by dominance of one or a few studies that measured a large number of responses, since when multiple responses were condensed into a single average within a trait and within a study, outcomes were very similar (Fig. \ref{FigA5}).

We were pleasantly surprised to find that correlations had been reported for at least two different stages across 18 traits out of the 20 targeted (Fig. \ref{Fig1}). We had thought it likely that for some traits correlations would be reported only for one size class, or only mixed across different sizes. These limitations applied only for LAR and NARmass, where studies appeared confined to juvenile stages. However, among the 20 traits studied we had enough data to test the effect of stage on the correlation between growth and traits only for the five for which we had predictions.

We found 112 studies published between between 1983 and 2014 that met our requirements. Climate zones spanned from temperate regions to  tropics (map in Fig. \ref{FigA2}). Different experiment types (laboratory, field, plantation or greenhouse), growth measurements and growth form (tree, woody or across growth form) were included. Number of species used for each correlation ranged from 2 to 300 species with average 31. Given this variety, it was not surprising to find wide ranges of r reported for all traits. For instance, r varied from -0.98 to 0.99 for SLA, and from -0.74 to 0.74 for WD (Fig. \ref{FigA3}). The growth form, the experiment types and the growth measurement were not markedly best predictors of the data (results shown in the Fig. \ref{FigA5}).

We met all the quality criteria required for meta-analysis in ecology (Table 2). We showed that the impact of publication bias is probably trivial in our meta-analyses (Appendix D, Fig. \ref{FigA7}). The extend of the heterogenity was high in our meta-analysis, and substantially explained by the plant stage for SLA and Seedmass as expected (Appendix D). Finally, temporal change in effect size did not jeopardizing the stability of our conclusions for all the traits (Appendix D, Fig. \ref{FigA9}).

@@

For seedlings our meta-analyses showed as expected a negative relationship between seed mass and relative growth rate (in mass or height or volume) that disappeared across plant stage. This negative relationship arises because seedlings from larger seeds are larger to begin with, as extensively discussed in the literature \citep[reviewed by][]{Turnbull:2012ew}. It applies only to RGR not to absolute growth rate, but for seedlings growth is most often reported as RGR (Appendix Fig. \ref{FigA8} d).

@@

\clearpage
\section*{Figures}

\begin{figure}[h!]
\centering
\includegraphics{output/Fig1.pdf}
\caption{Number of correlations between trait and growth for each plant stage recorded in our dataset. Shown here are 'raw dataset' (recorded both across and within species, under unstressed and stressed conditions). Bar colors: red for adult, orange for sapling, grey for juvenile. Aarea (or Amass) = rate of $CO_{2}$ assimilation per unit leaf area (or leaf mass) = maximum photosynthetic rate, Hmax= asymptotic maximum height,  Ks =stem hydraulic conductance, LA = leaf area, LMR = leaf mass ratio, NARarea (or NARmass) = net mass assimilation rate per unit of leaf area (or leaf mass), Nmass (or Narea) = nitrogen content per unit of leaf mass (or leaf area), Pmass = phosphorus content per unit of leaf mass, SA/LA = sapwood area per leaf area, SLA = specific leaf area, thickness = leaf thickness, Vessel size = stem conduit size, vessel density = stem conduit density, WD = wood density}
\label{Fig1}
\end{figure}

\begin{figure}[h!]
\centering
\includegraphics{output/Fig2.pdf}
\caption{Range of plant size in \textbf{a)} age, \textbf{b)} height and \textbf{c)} diameter and the plant stage assigned by the authors of the articles used in the meta-analyses. Each segment represents the plant size range (max and min values) of the species used to measure the correlation between growth and traits in each article. ID is the reference number of the article in our dataset. The colors represents the stage information given by the authors (green: seedling, blue: juvenile, orange: sapling, red: adult, grey: unspecified by the authors), the vertical line represents the most likely limits between stage categories, the horizontal line styles corresponds to the species growth form used in each study (solid line: exclusively tree, dashed line: woody species , dotted line: tree mixed with another growth form). Shown here are the "raw dataset" (recorded both across and within species, and under unstressed and stressed conditions).}
\label{Fig2}
\end{figure}

\begin{figure}[h!]
\centering
\includegraphics{output/Fig3.pdf}
\caption{The relationship between growth rates and SLA or seed mass changes with plant stage. Plots show the effect size (+ 95\%CI) of the correlation across species between growth rate and five functional traits for plant stage comparisons. \textbf{a)} Specific leaf area (SLA), \textbf{b)} Wood density (WD), \textbf{c)} Asymptotic height (Hmax) \textbf{d)} Seed mass and \textbf{e)} Assimilation rate of CO2 (Aarea). Effect size is a standardized measure of the deviation of correlation coefficient from zero (Fisher's z-transformed). Effects are significant if confidence intervals (95\%CI) for a given stage do not overlap with zero. Shown here are 'across species dataset' (recorded only across species, and under both unstressed and stressed conditions). Black point: unstressed conditions, grey points: all the data. 'n' indicates number of correlations contributing to each effect. Likelihood Ratio Tests (LRT, for unstressed conditions only) compared maximum-likelihood fits between a pooled model and one separated by stage. A significant p-value means the stage model fitted the data better than the null model. 
}
\label{Fig3}
\end{figure}

\clearpage
\section*{Table}
\setcounter{table}{0}
\begin{table}[h!]
\centering
\caption{\textbf{Hypothesised effects of traits on key elements of plant function determining growth rate.} Arrows indicate the effect an increase in trait value would have on the main elements of growth eqs. \ref{eq:eq0}-\ref{eq:eq4} in Text box 1:  Leaf deployment per mass ($\frac{\textrm{d}a_\textrm{l}}{\textrm{d}m_\textrm{t}}$), Allocation to vegetative growth ($\frac{\textrm{d}m_\textrm{t}}{\textrm{d}b}$), Net live biomass production ($\frac{\textrm{d}b}{\textrm{d}t}$), Sapwood turnover ($k_s a_s$). SLA: Specific leaf area, WD: Wood density, Hmax: Asymptotic maximum height, SM: seed mass, Aarea: Maximum photosynthetic rate. For further details, see Text Box 1.}
\vspace{1cm}
  \begin{tabular}{l p{3cm} p{2cm}p{4cm} p{2cm}}
  & Leaf deployment \newline per mass
  & Allocation to vegetative growth
  & Net live biomass \newline production
  & Sapwood turnover\\
  & & & &\\\hline
  & & & & \\
Increase in : & & & & \\ 
  SLA &$\nearrow$ &$\_$ & $\searrow$ \newline (via $\nearrow$ of tissue turnover) &$\_$ \\
  WD & $\searrow$ &$\_$  &$\_$ &$\_$ \\
  Hmax &$\_$ &$\nearrow$ &$\_$ &$\_$ \\
  SM &$\_$&$\_$&$\_$& $\_$\\ 
  Aarea &$\_$ &$\_$ & $\nearrow$ &$\_$ \\
\hline
  \end{tabular}
  \end{table}
\label{tab:trade-offs}

\clearpage
\begin{table}[h!]
\centering
\caption{\textbf{Checklist of quality criteria for meta-analysis in ecology} developed by Koricheva and Gurevitch (2014) }
\label{Table 2}
\vspace{0.5cm}
\begin{tabular}{p{3cm} p{1cm} p{10cm} p{3cm}}
  \hline
  Quality criteria & Check & Explanations & Addressed in \\
  \hline
  1. Reporting full details of bibliographic searches & yes	& Google Scholar database was searched using the range of Boolean search terms ((''SLA'' OR ''LMA'' OR ''specific leaf area'' OR ''leaf    mass area'' OR ''wood density'' OR ''WD'' OR ''wood specific gravity'' OR ''stem density'' OR ''stem specific density'') AND (''growth rate'' OR ''relative growth rate'')) AND (tree* OR woody*), in english. & "Systematic search", Table A1 \\
  2. Reporting inclusion/exclusion criteria & yes & Studies included report i) correlation coefficient between growth and trait, ii) the sample size/ number of species, and iii) the size/stage of plants. Studies excluded report correlation within species	& "Systematic search" \\
  3. Weighting effect sizes by study precision & yes & Effects sizes were weighted by number of species. The question addressed in our study concerns the correlation between growth and trait across species, thus the number of species is a better proxy of a study quality than sample size. &	 "Systematic search" and  "Effect size calculation" \\
 4. Specifying meta-analytical model &	yes	& Mixed-effect model with a study ID as a random variation, and stage as a fixed effect, m1 ~lmer (stage + (1|ID)) &	"Statistical analyses", Fig3 \\
 5. Quantifying heterogeneity in effect sizes &	yes &	We estimated the extend of the heterogeneity using the statistics $I^{2}$ (Santos and Nakagawa, 2012) for each trait. &	Appendix D \\
 6. Exploring causes of heterogeneity &	yes &	We explored two causes of variability: the stage and the year of publication & Appendix D, Fig3 \\
 7. Multifactorial analysis of explanatory variable	& yes & We focused on the "plant stage" as the main exploratory variable. Yet, the stage variable may be confunded with other explanatory moderators (i.e. growth measurement, experiment type or growth form). Thus, we checked if no one of them would not consistenly be a strongest predictor of the data. We did not test the multiple variables in a hierarchical fashion, because of the limited size of our dataset. & FigA4, FigA6  \\
  8. Testing for publication bias &	yes	& We used funnel plots for a visual assessement. We  used the "trim and fill" method to estimate the number of studies missing (Duval and Tweedie 2000). We also estimated the Rosenberg’s fail-safe number (Rosenberg 2005). 	&  Appendix D, FigA7\\
  9. Sensitivity analysis &	yes &	We tested how robust are our results to the decision made in the analysis. We calculated effect size using 3 differents dataset: i) an  "ideal" dataset, focusing on unstressed conditions exclusively, ii) a "complete dataset" reporting all conditions, and iii) a "reduced dataset" where multiple responses were condensed into a single average within a trait and within a study. &	Fig3, FigA5 \\
  10.Exploring temporal changes in effect size &	yes	& There may be a confounding effect between the publication year and the stage. Thus, we explored the publication effect on the effect size within each stage for traits where a stage effect had been highlighted (SLA and Seedmass). & Appendix D, FigA9 \\
  11. Controlling for phylogeny &	no &	It was not possible to control by the phylogeny since the number of species used to measured each coeffcient of correlation ranged from 2 to 300 species. &	\\
  12. Specifying the software used &	yes	& R softward, code available on Github. & "Availability of code and data" \\
 13. Providing reference list of primary studies included in the analysis &	yes &	Available on Github. The Bibtex file "meta-analyses.bib" corresponds to all the articles included in the meta-analyses, "read.bib" corresponds to all the articles (included and excluded) inspected for the meta-analyses.	& "Availability of code and data" \\
  14. Providing the data set used for meta-analysis  &	yes &	A copy of the complete data set is available on Github.	& "Availability of code and data" \\
   \hline
\end{tabular}
\end{table}

\clearpage
\section*{Text Box 1} \label{sec:growth}

@@

\clearpage
\begin{appendices}\label{sec:appendices}
\renewcommand{\thefigure}{S\arabic{figure}}
\renewcommand{\thetable}{S\arabic{table}}

@@

\clearpage
\section{Supporting tables}\label{app:supp_info_tables}
